---
title: "Crop productivity analysis"
---

## Packages

```{r}
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)
library(MASS) #boxcox transformation
library(lme4)         #mixed models
library(lmerTest)     #mixed models
library(broom.mixed)  #mixed models
library(emmeans)
```

## Loading Data

```{r}
yield <- read.csv(here::here("00-data", "b-prepared", 
                             "20250530_SCM-yield.csv"), 
                      header = TRUE)

spad <- read.csv(here::here("00-data", "b-prepared", 
                            "20250716_SCM-SPAD.csv"), 
                     header = TRUE)

```

## Format Data

```{r}
#separate edge data


spad <- spad %>%
        subset(block != 5) %>% 
        dplyr::select(-average_reading) %>%
        mutate(date_collected = lubridate::as_date(date_collected, format = "%m/%d/%Y"),
               year = lubridate::isoyear(ymd(date_collected))) %>%
        pivot_longer(cols = c("reading_1", "reading_2", "reading_3"),
                     names_to = "reading",
                     values_to = "value") %>%
        mutate_at(c('sampling','block', 'plot', 'treatment', 'crop', "year"), as.factor)
```

## Yield

### Visualization

```{r}


```

### Normality

```{r}

```

### ANOVA

```{r}

```

### Linear Mixed Model

```{r}

```

## Foliar Chlorophyll

### Visualization

Most relevant visual for my analysis thus far
```{r}

#sampling combined, visualized by crop/treatment for each year
ggplot(spad, aes(x = crop, y = value, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

```


Other visualizations
```{r}

#visualized each year separately, by sampling for each crop/treatment
ggplot(subset(spad, year == 2023), aes(x = sampling, y = value, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)
ggplot(subset(spad, year == 2024), aes(x = sampling, y = value, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#years combined, visualized by sampling for each crop/treatment. REVISE THIS LATER, since there are a different number of sampling dates in each year
ggplot(spad, aes(x = sampling, y = value, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#sampling and year combined, visualized by crop/treatment. 
ggplot(spad, aes(x = crop, y = value, fill = treatment)) + geom_boxplot()

#crops combined, year visualized by sampling/treatment
ggplot(spad, aes(x = sampling, y = value, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

#treatments combined, year visualized by sampling/crop
ggplot(spad, aes(x = sampling, y = value, fill = crop)) + geom_boxplot() + facet_wrap(~year)

#crops and sampling combined, visualized by treatment for each year
ggplot(spad, aes(x = year, y = value, fill = treatment)) + geom_boxplot()

#treatments and sampling combined, visualized by crop for each year
ggplot(spad, aes(x = year, y = value, fill = crop)) + geom_boxplot()

# just treatments
ggplot(spad, aes(x = treatment, y = value)) + geom_boxplot()

#just crops
ggplot(spad, aes(x = crop, y = value)) + geom_boxplot()
```

### Normality
At what point do we mark values as NA because the SPAD just can't go that high?

```{r}
hist(spad$value)
skewness(spad$value, na.rm = TRUE)
qqnorm(spad$value, main='Normal')
qqline(spad$value)
shapiro.test(spad$value)
```

#### boxcox transformation
This gets me closer to a normal distribution
```{r}
bc <- boxcox(lm(spad$value ~ 1))
lambda <- bc$x[which.max(bc$y)]
spad$bc_value <- (spad$value ^ lambda -1)/lambda

hist(spad$bc_value)
skewness(spad$bc_value, na.rm = TRUE)
qqnorm(spad$bc_value, main='Normal')
qqline(spad$bc_value)
shapiro.test(spad$bc_value)

```

### Kruskal-Wallis

```{r}
#across the whole data set, there is a difference between crops
kruskal.test(value ~ year, data = spad)
kruskal.test(value ~ treatment, data = spad)
kruskal.test(value ~ crop, data = spad)

#within each crop, there is a difference between years
kruskal.test(value ~ year, data = subset(spad, crop == "AS"))
kruskal.test(value ~ year, data = subset(spad, crop == "DB"))
kruskal.test(value ~ year, data = subset(spad, crop == "FC"))

#within each crop, there is only a difference between treatments for AS
kruskal.test(value ~ treatment, data = subset(spad, crop == "AS"))
kruskal.test(value ~ treatment, data = subset(spad, crop == "DB"))
kruskal.test(value ~ treatment, data = subset(spad, crop == "FC"))


#within a crop-year combination, there is a significant difference between treatments for AS in 2023, between DB in 2024, and between FC in 2024
kruskal.test(value ~ treatment, data = subset(spad, crop == "AS" & year == "2023"))
kruskal.test(value ~ treatment, data = subset(spad, crop == "AS" & year == "2024"))

kruskal.test(value ~ treatment, data = subset(spad, crop == "DB" & year == "2023"))
kruskal.test(value ~ treatment, data = subset(spad, crop == "DB" & year == "2024"))

kruskal.test(value ~ treatment, data = subset(spad, crop == "DB" & year == "2023"))
kruskal.test(value ~ treatment, data = subset(spad, crop == "DB" & year == "2024"))

```

### Linear Mixed Model

```{r}

```
