---
title: "Weed Biomass and Labor"
---

## Packages

```{r}
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)

```

## Loading Data

```{r}
weed_biomass <- read.csv(here::here("00-data", "b-prepared", 
                                 "20250714_SCM-weed biomass.csv"), 
                      header = TRUE)

weed_labor <- read.csv(here::here("00-data", "b-prepared", 
                                  "20250714_SCM-weed labor.csv"), 
                     header = TRUE)

```

## Format Data

```{r}
weed_biomass <- weed_biomass %>%
                subset(block != 5) %>%
                mutate(date_collected = lubridate::as_date(date_collected, format = "%m/%d/%Y"),
                       year = lubridate::isoyear(ymd(date_collected)),
                       od_g = ifelse(od_biomass_calculated_g < 0, 0, od_biomass_calculated_g)) %>%
                mutate_at(c('weeding_point', 'block', 'plot', 'treatment', 'crop'), as.factor)

weed_labor <- weed_labor %>%
              subset(block != 5) %>% 
              mutate(date_collected = lubridate::as_date(date_collected, format = "%m/%d/%Y"),
                     year = lubridate::isoyear(ymd(date_collected))) %>%
              mutate_at(c('weeding_point', 'block', 'plot', 'treatment', 'crop', "collected_by", "year"), as.factor) %>%
              dplyr::select(-c(start.time, end.time, total.time))
```

## Weed Biomass

### Visualization

Most representative of statistical analysis
```{r}
#crops combined, year visualized by weed collection/treatment
ggplot(weed_biomass, aes(x = weeding_point, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

```

Additional visualizations
```{r}
#visualized each year separately, by weed collection for each crop/treatment
ggplot(subset(weed_biomass, year == 2023), aes(x = weeding_point, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)
ggplot(subset(weed_biomass, year == 2024), aes(x = weeding_point, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#years combined, visualized by weed collection for each crop/treatment
ggplot(weed_biomass, aes(x = weeding_point, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#weeding point combined, visualized by crop/treatment for each year
ggplot(weed_biomass, aes(x = crop, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

#weeding point and year combined, visualized by crop/treatment
ggplot(weed_biomass, aes(x = crop, y = od_g, fill = treatment)) + geom_boxplot()

#crops and weeding point combined, visualized by treatment for each year
ggplot(weed_biomass, aes(x = as.factor(year), y = od_g, fill = treatment)) + geom_boxplot()

# just treatments
ggplot(weed_biomass, aes(x = treatment, y = od_g)) + geom_boxplot()

```

### Normality
Is it possible to transform this to a normal distribution, and if not, how do I analyze it?

```{r}
library(moments) #skewness function

hist(weed_biomass$od_g)
skewness(weed_biomass$od_g, na.rm = TRUE)
qqnorm(weed_biomass$od_g, main='Normal')
qqline(weed_biomass$od_g)
shapiro.test(weed_biomass$od_g)

bc <- boxcox(lm((weed_biomass$od_g + 1) ~ 1))
lambda <- bc$x[which.max(bc$y)]
weed_biomass$bc_od <- ((weed_biomass$od_g + 1) ^ lambda -1)/lambda

```


### One way tests

```{r}
library(FSA)          # <3>
kruskal.test(od_g ~ year, weed_biomass) # p < 0.001, so analyzing years separately

kruskal.test(od_g ~ weeding_point, subset(weed_biomass, year == "2023")) # p < 0.001, so analyzing each weeding point separately
kruskal.test(od_g ~ weeding_point, subset(weed_biomass, year == "2024")) # p < 0.001, so analyzing each weeding point separately
```

#### 2023

##### Weeding Day 1
When crops are combined, there were significantly more weeds in bare soil compared to cardboard mulch (p < 0.002). Weed biomass under straw mulch was not significantly different from bare soil or cardboard.
```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2023" & weeding_point == "1")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "1"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "1")) 

```

##### Weeding Day 2
When crops are combined, there were significantly less weed biomass under cardboard mulch compared to bare soil (p = 0.007) and straw mulch (p = 0.02). 
```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2023" & weeding_point == "2")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "2"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "2")) 

```

##### Weeding Day 3
When we reached the third weeding day, there was a significant difference in weed biomass between crops was observed (p = 0.01). For strawberries, weed biomass under cardboard mulch was significantly less than bare soil (p = 0.03). All other crop-treatment combinations were not significantly different from each other. 

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2023" & weeding_point == "3")) # p = 0.01, so crops are analyzed individually (plus I also combine for consistency with previous days)

#Strawberries 
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "AS"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "AS")) 

#Dragon tongue beans 
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "DB"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "DB")) 

#Collards
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "FC"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "FC")) 

#crops combined
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3")) 

```
##### Weeding Day 4
Mulch treatment did not impact weed biomass during our fourth collection point (p = 0.4).

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2023" & weeding_point == "4")) # p = 0.4, so crops were combined
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "4"))

```

#### 2024

##### Weeding Day 1
When crops were combined, weed biomass was significantly higher in bare soil compared to both mulch treatments (p < 0.001). There was not, however, a difference in biomass between cardboard and straw mulch (p = 0.5).

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2024" & weeding_point == "1")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "1"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "1")) 

```

##### Weeding Day 2
When crops were combined, weed biomass was significantly higher in bare soil compared to cardboard (p < 0.001) and straw (p = 0.004) mulch. There was not, however, a difference in biomass between the two mulches (p = 0.6) 

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2024" & weeding_point == "2")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "2"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "2")) 

```

##### Weeding Day 3
When we reached the third weeding day, however, weed biomass in bare soil was only significantly higher than cardboard mulch (p = 0.009).

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2024" & weeding_point == "3")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "3"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "3")) 

```

##### Weeding Day 4
During our final weeding day, weed biomass in bare soil remained significantly higher than for cardboard mulch (p = 0.003). However, now, weed biomass in straw mulch was also significantly higher than for cardboard (p < 0.001)

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2024" & weeding_point == "4")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "4"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "4")) 

```

### Linear Mixed Model

```{r}

```

## Weed Labor 

### checking for impact of person weeding
I didn't actually do this first, but I found what was likely a mis-recorded time that I want to remove from analysis

```{r}
# weeding time visualized for each person  
ggplot(subset(weed_labor, year == "2023"), aes(x = treatment, y = time_min, color = collected_by)) + geom_boxplot() 
ggplot(subset(weed_labor, year == "2024" & treatment == "CM"), aes(x = weeding_point, y = time_min, color = collected_by)) + geom_boxplot() 

plots_by_person <- weed_labor %>%
                   group_by(treatment, collected_by) %>%
                   summarize(count = n(),
                             avg_time = mean(time_min)) %>%
                   ungroup()
plots_by_person

# I checked for interactions with who did the weeding, and I didn't see any concerning ones. (I went through the TukeyHSD results and checked each comparison where two people had weeded the same mulch treatment. Only one combination was signicant for SM: Trick and Nora, and I know from our notes and my memories that Trick just happened to tackle several plots that were especially weedy)
aov <- aov(bc_time ~ treatment*collected_by, data = subset(weed_labor, year == "2023"))
summary(aov)
TukeyHSD(aov)

# Checking for differences in weeding time for different people in 2024. I went through the TukeyHSD results and checked each comparison where two people had weeded the same mulch treatment. Morgan Smith was significantly different than MH (slightly: 0.07), NP (0.006), PM (0.001), and RD (0.001) for cardboard mulch, and slight different from PM for straw mulch (0.06). When compared to the ggplot, I noted that NP weeded earlier in the season (and thus would likely have had longer weeding times). MH had times that were in a similar range with other people who weeded the same weeks as they did. Comparing with biomass, PM was in charge of a plot with A LOT of weed biomass during the second weed collection. RD, however, weeded 1-E during the second collection point, and there was only 0.99 g of dried biomass but it took them 68 minutes. I'm assuming that this was an error recording time and, consequently, removed.
aov <- aov(bc_time ~ treatment*collected_by, data = subset(weed_labor, year == "2024"))
summary(aov)
TukeyHSD(aov)

#use this if you actually want to see the whole output in the previous tukey hsd
#options(max.print = 2150)

ggplot(subset(weed_labor, year == "2024" & treatment == "CM"), aes(x = weeding_point, y = time_min, color = collected_by)) + geom_boxplot() 


```


```{r}
#creating a copy that includes the suspected bad data (weed_labor_out) so I can check whether it impacts analysis
weed_labor_out <- weed_labor
weed_labor$time_min[weed_labor$weeding_point == "2" & weed_labor$block == "1" & weed_labor$plot == "E" & weed_labor$year == "2024"] <- NA

```

### Visualization
This makes me have suspicions about whether they were using hand tools to weed in 2024.
```{r}
#visualized each year separately, by weed collection for each crop/treatment
ggplot(subset(weed_labor, year == 2023), aes(x = weeding_point, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)
ggplot(subset(weed_labor, year == 2024), aes(x = weeding_point, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#years combined, visualized by weed collection for each crop/treatment
ggplot(weed_labor, aes(x = weeding_point, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#weeding point combined, visualized by crop/treatment for each year
ggplot(weed_labor, aes(x = crop, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

#weeding point and year combined, visualized by crop/treatment
ggplot(weed_labor, aes(x = crop, y = time_min, fill = treatment)) + geom_boxplot()

#crops combined, year visualized by weed collection/treatment
ggplot(weed_labor, aes(x = weeding_point, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

#crops and weeding point combined, visualized by treatment for each year
ggplot(weed_labor, aes(x = as.factor(year), y = time_min, fill = treatment)) + geom_boxplot()

# just treatments
ggplot(weed_labor, aes(x = treatment, y = time_min)) + geom_boxplot()


```

### Normality

```{r}
hist(weed_labor$time_min)
skewness(weed_labor$time_min, na.rm = TRUE)
qqnorm(weed_labor$time_min, main='Normal')
qqline(weed_labor$time_min)
shapiro.test(weed_labor$time_min)
```
# log transformation
```{r}
weed_labor$log_time <- log10(weed_labor$time_min)
hist(weed_labor$log_time)
skewness(weed_labor$log_time, na.rm = TRUE)
qqnorm(weed_labor$log_time, main='Normal')
qqline(weed_labor$log_time)
shapiro.test(weed_labor$log_time)

```
# boxcox transformation
This gets me slightly closer to a normal distribution
```{r}
bc <- boxcox(lm(weed_labor$time_min ~ 1))
lambda <- bc$x[which.max(bc$y)]
weed_labor$bc_time <- (weed_labor$time_min ^ lambda -1)/lambda

hist(weed_labor$bc_time)
skewness(weed_labor$bc_time, na.rm = TRUE)
qqnorm(weed_labor$bc_time, main='Normal')
qqline(weed_labor$bc_time)
shapiro.test(weed_labor$bc_time)

```

### ANOVA

#### Initial
```{r}
#This demonstrates treatment, crop, and year have a significant effect == but only treatment year has a significant interaction effect. 
aov <- aov(bc_time ~ treatment*crop*year, data = weed_labor)
summary(aov)
TukeyHSD(aov)

#Based on broader results above, I'm analyzing each year separtely. 
#In this first ANOVA for 2023, only treatment is significant
aov <- aov(bc_time ~ treatment*crop, data = subset(weed_labor, year == "2023"))
summary(aov)
TukeyHSD(aov)

# In 2023, bare soil and cardboard mulch required similar time to weed, but straw mulch was significantly different from both
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2023"))
summary(aov)
TukeyHSD(aov)


# In 2024, there was once again no interaction between treatment and crop. Crop had a potential significant impact on weeding time (0.05) in the ANOVA, but not the Tukey HSD
aov <- aov(bc_time ~ treatment*crop, data = subset(weed_labor, year == "2024"))
summary(aov)
TukeyHSD(aov)

# In 2024, we saw a different trend. This year, cardboard mulch required significantly less weeding time than both bare soil and straw mulch (and there was no significant difference between bare soil and straw).
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2024"))
summary(aov)
TukeyHSD(aov)

# Still, I'm curious about if we analyze the years together. We cardboard mulch is significantly different from bare soil and straw. Straw mulch and bare soil are *nearly* significantly different (0.06)
aov <- aov(bc_time ~ treatment, data = weed_labor)
summary(aov)
TukeyHSD(aov)

```
#### For each weeding day (from LMM results)




### Linear Mixed Model

```{r}
model <- lmer(bc_time ~ treatment*crop*year + (1 | weeding_point/block), data = weed_labor)
AIC(model)
summary(model)

#extract random effects and residual values
re <- ranef(model)
resid <- augment(model)

#generate qqplot
plot(re)

#generate plot of fitted vs. residual values
ggplot(resid, aes(x = .fitted, y = .resid)) + geom_point() + stat_smooth(se = F)

#ranked likelihood test
ranova(model)

#emmeans(model, pairwise ~ treatment*year)

```
