---
title: "Weed Biomass and Labor"
---

## Packages

```{r}
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)
library(MASS) #boxcox transformation
library(lme4)         #mixed models
library(lmerTest)     #mixed models
library(broom.mixed)  #mixed models
library(emmeans)

```

## Loading Data

```{r}
weed_biomass <- read.csv(here::here("00-data", "b-prepared", 
                                 "20250714_SCM-weed biomass.csv"), 
                      header = TRUE)

weed_labor <- read.csv(here::here("00-data", "b-prepared", 
                                  "20250714_SCM-weed labor.csv"), 
                     header = TRUE)

```

## Format Data

```{r}
weed_biomass <- weed_biomass %>%
                subset(block != 5) %>%
                mutate(date_collected = lubridate::as_date(date_collected, format = "%m/%d/%Y"),
                       year = lubridate::isoyear(ymd(date_collected)),
                       od_g = ifelse(od_biomass_calculated_g < 0, 0, od_biomass_calculated_g)) %>%
                mutate_at(c('weeding_point', 'block', 'plot', 'treatment', 'crop'), as.factor)

weed_labor <- weed_labor %>%
              subset(block != 5) %>% 
              mutate(date_collected = lubridate::as_date(date_collected, format = "%m/%d/%Y"),
                     year = lubridate::isoyear(ymd(date_collected))) %>%
              mutate_at(c('weeding_point', 'block', 'plot', 'treatment', 'crop', "collected_by", "year"), as.factor) %>%
              dplyr::select(-c(start.time, end.time, total.time))
```

## Weed Biomass

### Visualization

Most representative of statistical analysis
```{r}
#crops combined, year visualized by weed collection/treatment
ggplot(weed_biomass, aes(x = weeding_point, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

```

Additional visualizations
```{r}
#visualized each year separately, by weed collection for each crop/treatment
ggplot(subset(weed_biomass, year == 2023), aes(x = weeding_point, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)
ggplot(subset(weed_biomass, year == 2024), aes(x = weeding_point, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#years combined, visualized by weed collection for each crop/treatment
ggplot(weed_biomass, aes(x = weeding_point, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#weeding point combined, visualized by crop/treatment for each year
ggplot(weed_biomass, aes(x = crop, y = od_g, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

#weeding point and year combined, visualized by crop/treatment
ggplot(weed_biomass, aes(x = crop, y = od_g, fill = treatment)) + geom_boxplot()

#crops and weeding point combined, visualized by treatment for each year
ggplot(weed_biomass, aes(x = as.factor(year), y = od_g, fill = treatment)) + geom_boxplot()

# just treatments
ggplot(weed_biomass, aes(x = treatment, y = od_g)) + geom_boxplot()

```

### Normality
Is it possible to transform this to a normal distribution, and if not, how do I analyze it?

```{r}
library(moments) #skewness function

hist(weed_biomass$od_g)
skewness(weed_biomass$od_g, na.rm = TRUE)
qqnorm(weed_biomass$od_g, main='Normal')
qqline(weed_biomass$od_g)
shapiro.test(weed_biomass$od_g)

bc <- boxcox(lm((weed_biomass$od_g + 1) ~ 1))
lambda <- bc$x[which.max(bc$y)]
weed_biomass$bc_od <- ((weed_biomass$od_g + 1) ^ lambda -1)/lambda

```


### One way tests

```{r}
library(FSA)          # <3>
kruskal.test(od_g ~ year, weed_biomass) # p < 0.001, so analyzing years separately

kruskal.test(od_g ~ weeding_point, subset(weed_biomass, year == "2023")) # p < 0.001, so analyzing each weeding point separately
kruskal.test(od_g ~ weeding_point, subset(weed_biomass, year == "2024")) # p < 0.001, so analyzing each weeding point separately
```

#### 2023

##### Weeding Day 1
When crops are combined, there were significantly more weeds in bare soil compared to cardboard mulch (p < 0.002). Weed biomass under straw mulch was not significantly different from bare soil or cardboard.
```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2023" & weeding_point == "1")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "1"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "1")) 

```

##### Weeding Day 2
When crops are combined, there were significantly less weed biomass under cardboard mulch compared to bare soil (p = 0.007) and straw mulch (p = 0.02). 
```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2023" & weeding_point == "2")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "2"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "2")) 

```

##### Weeding Day 3
When we reached the third weeding day, there was a significant difference in weed biomass between crops was observed (p = 0.01). For strawberries, weed biomass under cardboard mulch was significantly less than bare soil (p = 0.03). All other crop-treatment combinations were not significantly different from each other. 

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2023" & weeding_point == "3")) # p = 0.01, so crops are analyzed individually (plus I also combine for consistency with previous days)

#Strawberries 
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "AS"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "AS")) 

#Dragon tongue beans 
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "DB"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "DB")) 

#Collards
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "FC"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3" & crop == "FC")) 

#crops combined
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "3")) 

```
##### Weeding Day 4
Mulch treatment did not impact weed biomass during our fourth collection point (p = 0.4).

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2023" & weeding_point == "4")) # p = 0.4, so crops were combined
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2023" & weeding_point == "4"))

```

#### 2024

##### Weeding Day 1
When crops were combined, weed biomass was significantly higher in bare soil compared to both mulch treatments (p < 0.001). There was not, however, a difference in biomass between cardboard and straw mulch (p = 0.5).

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2024" & weeding_point == "1")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "1"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "1")) 

```

##### Weeding Day 2
When crops were combined, weed biomass was significantly higher in bare soil compared to cardboard (p < 0.001) and straw (p = 0.004) mulch. There was not, however, a difference in biomass between the two mulches (p = 0.6) 

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2024" & weeding_point == "2")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "2"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "2")) 

```

##### Weeding Day 3
When we reached the third weeding day, however, weed biomass in bare soil was only significantly higher than cardboard mulch (p = 0.009).

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2024" & weeding_point == "3")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "3"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "3")) 

```

##### Weeding Day 4
During our final weeding day, weed biomass in bare soil remained significantly higher than for cardboard mulch (p = 0.003). However, now, weed biomass in straw mulch was also significantly higher than for cardboard (p < 0.001)

```{r}
kruskal.test(od_g ~ crop, subset(weed_biomass, year == "2024" & weeding_point == "4")) # p > 0.05, so combining crops to look at treatments
kruskal.test(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "4"))
dunnTest(od_g ~ treatment, subset(weed_biomass, year == "2024" & weeding_point == "4")) 

```

### Linear Mixed Model

```{r}

```

## Weed Labor 

### checking for impact of person weeding
I didn't actually do this first, but I found what was likely a mis-recorded time that I want to remove from analysis

```{r}
# weeding time visualized for each person  
ggplot(subset(weed_labor, year == "2023"), aes(x = treatment, y = time_min, color = collected_by)) + geom_boxplot() 
ggplot(subset(weed_labor, year == "2024" & treatment == "CM"), aes(x = weeding_point, y = time_min, color = collected_by)) + geom_boxplot() 

plots_by_person <- weed_labor %>%
                   group_by(treatment, collected_by) %>%
                   summarize(count = n(),
                             avg_time = mean(time_min)) %>%
                   ungroup()
plots_by_person

# I checked for interactions with who did the weeding, and I didn't see any concerning ones. (I went through the TukeyHSD results and checked each comparison where two people had weeded the same mulch treatment. Only one combination was signicant for SM: Trick and Nora, and I know from our notes and my memories that Trick just happened to tackle several plots that were especially weedy)
aov <- aov(bc_time ~ treatment*collected_by, data = subset(weed_labor, year == "2023"))
summary(aov)
TukeyHSD(aov)

# Checking for differences in weeding time for different people in 2024. I went through the TukeyHSD results and checked each comparison where two people had weeded the same mulch treatment. Morgan Smith was significantly different than MH (slightly: 0.07), NP (0.006), PM (0.001), and RD (0.001) for cardboard mulch, and slight different from PM for straw mulch (0.06). When compared to the ggplot, I noted that NP weeded earlier in the season (and thus would likely have had longer weeding times). MH had times that were in a similar range with other people who weeded the same weeks as they did. Comparing with biomass, PM was in charge of a plot with A LOT of weed biomass during the second weed collection. RD, however, weeded 1-E during the second collection point, and there was only 0.99 g of dried biomass but it took them 68 minutes. I'm assuming that this was an error recording time and, consequently, removed.
aov <- aov(bc_time ~ treatment*collected_by, data = subset(weed_labor, year == "2024"))
summary(aov)
TukeyHSD(aov)

#use this if you actually want to see the whole output in the previous tukey hsd
#options(max.print = 2150)

ggplot(subset(weed_labor, year == "2024" & treatment == "CM"), aes(x = weeding_point, y = time_min, color = collected_by)) + geom_boxplot() 


```


```{r}
#creating a copy that includes the suspected bad data (weed_labor_out) so I can check whether it impacts analysis
weed_labor_out <- weed_labor
weed_labor$time_min[weed_labor$weeding_point == "2" & weed_labor$block == "1" & weed_labor$plot == "E" & weed_labor$year == "2024"] <- NA

```

### Visualization
This makes me have suspicions about whether they were using hand tools to weed in 2024.
```{r}
#visualized each year separately, by weed collection for each crop/treatment
ggplot(subset(weed_labor, year == 2023), aes(x = weeding_point, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)
ggplot(subset(weed_labor, year == 2024), aes(x = weeding_point, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#years combined, visualized by weed collection for each crop/treatment
ggplot(weed_labor, aes(x = weeding_point, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~crop)

#weeding point combined, visualized by crop/treatment for each year
ggplot(weed_labor, aes(x = crop, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

#weeding point and year combined, visualized by crop/treatment
ggplot(weed_labor, aes(x = crop, y = time_min, fill = treatment)) + geom_boxplot()

#crops combined, year visualized by weed collection/treatment
ggplot(weed_labor, aes(x = weeding_point, y = time_min, fill = treatment)) + geom_boxplot() + facet_wrap(~year)

#crops and weeding point combined, visualized by treatment for each year
ggplot(weed_labor, aes(x = as.factor(year), y = time_min, fill = treatment)) + geom_boxplot()

# just treatments
ggplot(weed_labor, aes(x = treatment, y = time_min)) + geom_boxplot()


```

### Normality

```{r}
hist(weed_labor$time_min)
skewness(weed_labor$time_min, na.rm = TRUE)
qqnorm(weed_labor$time_min, main='Normal')
qqline(weed_labor$time_min)
shapiro.test(weed_labor$time_min)
```
# log transformation
```{r}
weed_labor$log_time <- log10(weed_labor$time_min)
hist(weed_labor$log_time)
skewness(weed_labor$log_time, na.rm = TRUE)
qqnorm(weed_labor$log_time, main='Normal')
qqline(weed_labor$log_time)
shapiro.test(weed_labor$log_time)

```
# boxcox transformation
This gets me slightly closer to a normal distribution
```{r}
bc <- boxcox(lm(weed_labor$time_min ~ 1))
lambda <- bc$x[which.max(bc$y)]
weed_labor$bc_time <- (weed_labor$time_min ^ lambda -1)/lambda

hist(weed_labor$bc_time)
skewness(weed_labor$bc_time, na.rm = TRUE)
qqnorm(weed_labor$bc_time, main='Normal')
qqline(weed_labor$bc_time)
shapiro.test(weed_labor$bc_time)

```

### ANOVA

#### Initial
```{r}
#This demonstrates treatment, crop, and year have a significant effect == but only treatment year has a significant interaction effect. 
aov <- aov(bc_time ~ treatment*crop*year, data = weed_labor)
summary(aov)
TukeyHSD(aov)

#Based on broader results above, I'm analyzing each year separtely. 
#In this first ANOVA for 2023, only treatment is significant
aov <- aov(bc_time ~ treatment*crop, data = subset(weed_labor, year == "2023"))
summary(aov)
TukeyHSD(aov)

# In 2023, both ANOVA and KW found that bare soil and cardboard mulch required similar time to weed, and straw mulch was significantly different from both
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2023"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2023"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2023"))

# In 2024, there was once again no interaction between treatment and crop. Crop had a potential significant impact on weeding time (0.05) in the ANOVA, but not the Tukey HSD
aov <- aov(bc_time ~ treatment*crop, data = subset(weed_labor, year == "2024"))
summary(aov)
TukeyHSD(aov)


# In 2024, from ANOVA and KW, we saw a different trend. This year, cardboard mulch required significantly less weeding time than both bare soil and straw mulch (and there was no significant difference between bare soil and straw).
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2024"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2024"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2024"))

# Still, I'm curious about if we analyze the years together. Using the ANOVA, cardboard mulch is significantly different from bare soil and straw. Straw mulch and bare soil are *nearly* significantly different (0.06). In the KW, all treatments are significantly different from each other
aov <- aov(bc_time ~ treatment, data = weed_labor)
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = weed_labor)
dunnTest(time_min ~ treatment, data = weed_labor)

```

#### 2023
note that I did check each week for interactions with crop. I also ran LMMs for each weeding day with block as the random effect; it did not significantly improve the model. 

##### Weeding Day 1
Straw mulch is significantly different from cardboard and bare soil. Cardboard is nearly significantly different (0.06). KW test shows slightly different results; only straw mulch and cardboard mulch are significantly different
```{r}
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "1"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "1"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "1"))

```

##### Weeding Day 2
Straw mulch is significantly different from cardboard and bare soil. 
```{r}
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "2"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "2"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "2"))

```

##### Weeding Day 3
In a one-way for treatment, straw mulch is significantly different from bare soil and cardboard mulch. 

In a one-way for crop, there were no significant differences. 

Looking at interactions, a significant difference between FC and DB appeared (0.04). The Tukey HSD indicates that:
BM: no significant difference between crops
SM: no significant difference between crops
CM: no significant difference between crops

AS: straw is significantly different from bare soil
DB: no significant difference between mulch treatments
FC: straw is significantly different from bare soil
```{r}
aov <- aov(bc_time ~ treatment+crop, data = subset(weed_labor, year == "2023" & weeding_point == "3"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "3"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "3"))


aov <- aov(bc_time ~ treatment*crop, data = subset(weed_labor, year == "2023" & weeding_point == "3"))
summary(aov)
TukeyHSD(aov)



```
##### Weeding Day 4
According to the ANOVA, all mulches are signifcantly different from each other. According to the KW, only straw mulch and cardboard mulch are significantly different (though the others are close)
```{r}

aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "4"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "4"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2023" & weeding_point == "4"))

```

#### 2024
note that I did check each weeding day for interactions with crop. I also ran LMMs for each weeding day with block as the random effect; it did not significantly improve the model.

##### Weeding Day 1
No significant difference between mulch treatments
```{r}
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "1"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "1"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "1"))

```
##### Weeding Day 2
According to the ANOVA, cardboard mulch was significantly different from bare soil and straw. Note that the LMM for block showed that block impacted results, though ultimately the relationship between treatments was the same.

According to the KW, though, cardboard was only significantly different from bare soil
```{r}
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "2"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "2"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "2"))

```

##### Weeding Day 3
cardboard wasn't quite significantly different from bare soil (0.06), though it becomes significant (0.04) if crop is added (though there are no significant differences between crops, nor are there any significant interactions)

In the KW, cardboard and straw are significantly different

```{r}
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "3"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "3"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "3"))

aov <- aov(bc_time ~ treatment*crop, data = subset(weed_labor, year == "2024" & weeding_point == "3"))
summary(aov)
TukeyHSD(aov)

```
##### Weeding Day 4
According to both ANOVA and KW, cardboard was significantly different from bare soil and straw. There was a significant difference for crops as well, between collards and strawberries/beans. Looking at interactions:

BM: no significant difference between crops
CM: no significant difference between crops (FC and DB nearly so, though)
SM: no significant difference between crops

AS: cardboard was significantly different from bare soil & straw
DB: no significant difference between mulches
FC: cardboard was significantly different from bare soil

There's a handful of other interactions for different crop/mulch combinations, but not taking the time to go through them right now

```{r}
aov <- aov(bc_time ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "4"))
summary(aov)
TukeyHSD(aov)

kruskal.test(time_min ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "4"))
dunnTest(time_min ~ treatment, data = subset(weed_labor, year == "2024" & weeding_point == "4"))

aov <- aov(bc_time ~ treatment*crop, data = subset(weed_labor, year == "2024" & weeding_point == "4"))
summary(aov)
TukeyHSD(aov)

```

### Linear Mixed Model

```{r}
model <- lmer(bc_time ~ treatment*crop + (1 | block), data = subset(weed_labor, year == "2024" & weeding_point == "4"))
AIC(model)
summary(model)

#extract random effects and residual values
re <- ranef(model)
resid <- augment(model)

#generate qqplot
plot(re)

#generate plot of fitted vs. residual values
ggplot(resid, aes(x = .fitted, y = .resid)) + geom_point() + stat_smooth(se = F)

#ranked likelihood test
ranova(model)

emmeans(model, pairwise ~ treatment)

```
