---
title: "Soil fertility"
---

## Packages

```{r}
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)

```

## Loading Data
```{r}
#will also need to add in 
soil_fert <- read.csv(here::here("00-data", "b-prepared", 
                                 "20250530_SCM-soil-fertility.csv"), 
                           header = TRUE)

bd <- read.csv(here::here("00-data", "b-prepared", 
                          "20250530_SCM-BD.csv"), 
                     header = TRUE)

```

## Format Data

```{r}

```

## Initial Soil Properties

```{r}


```

## Phosphorus

### Visualization

```{r}


```

### Normality
```{r}

```

### ANOVA
```{r}

```

### Linear Mixed Model
```{r}

```

## Potassium

### Visualization

```{r}


```

### Normality
```{r}

```

### ANOVA
```{r}

```

### Linear Mixed Model
```{r}

```


## Carbon

### Total Carbon

#### Visualization

```{r}


```

#### Normality
```{r}

```

#### ANOVA
```{r}

```

#### Linear Mixed Model
```{r}

```

### Bulk Density

#### Visualization

```{r}


```

#### Normality
```{r}

```

#### ANOVA
```{r}

```

#### Linear Mixed Model
```{r}

```

## Nitrogen

### Nitrate

#### Visualization

```{r}


```

#### Normality
```{r}

```

#### ANOVA
```{r}

```

#### Linear Mixed Model
```{r}

```

### Ammonium

#### Visualization

```{r}


```

#### Normality
```{r}

```

#### ANOVA
```{r}

```

#### Linear Mixed Model
```{r}

```

### Total Nitrogen

#### Visualization

```{r}


```

#### Normality
```{r}

```

#### ANOVA
```{r}

```

#### Linear Mixed Model
```{r}

```

### ACE

#### Visualization

```{r}


```

#### Normality
```{r}

```

#### ANOVA
```{r}

```

#### Linear Mixed Model
```{r}

```


## C:N Ratio

### Visualization

```{r}


```

### Normality
```{r}

```

### ANOVA
```{r}

```

### Linear Mixed Model
```{r}

```
